---
title: "00_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### Load the data 

```{r get-data-option-2, echo=TRUE}
load("/home/u034/mfourman/ccp_data_final.rda")
```

### Some data cleaning and sorting
```{r}
ccp_data %>%  head (10)
```



```{r}
ccp_data_labs <- ccp_data %>%
  drop_na(daily_lbdat) %>% 
  drop_na(cestdat) 

ccp_data_labs %>% 
  distinct(subjid, daily_lbdat) %>% 
  nrow()
```
#Sort the dates into the correct format
ccp_data_labs <- ccp_data_labs %>%
  mutate(cestdat = lubridate::ymd(cestdat),
         daily_lbdat = lubridate::ymd(daily_lbdat)) 

# get rid weird crazy dates from 1918
ccp_data_labs <- ccp_data_labs %>% 
  filter(daily_lbdat >= "2020-01-01") %>% 
  filter(daily_lbdat <= "2022-01-01") %>% 
  filter(cestdat >= "2020-01-01") %>% 
  filter(cestdat <= "2022-01-01")

#check that the max and min dates are sensible
ccp_data_labs %>%
  summarise(max_lab_date = max(daily_lbdat),
            min_lab_date = min(daily_lbdat),
            max_symptom_date = max(cestdat),
            min_symptom_date = min(cestdat))

ccp_data_labs <- ccp_data_labs %>% 
  rename(lymphocyte_count = daily_lymp_lborres) %>% 
  rename(neutrophil_count = daily_neutro_lborres) %>%
  rename(platelet_count = daily_plt_lborres)


# take the difference between the two dates to work out time to symptom onset
ccp_data_labs <- ccp_data_labs%>% 
  mutate(lab_day_since_onset = daily_lbdat - cestdat) %>% 
  mutate(nl_ratio = neutrophil_count/lymphocyte_count) 

ccp_data_labs %>% 
  select(subjid, daily_lbdat, cestdat, 
         daily_wbc_lborres, lymphocyte_count, neutrophil_count, 
         platelet_count) %>% 
  missing_glimpse()

ccp_data_labs %>% 
  distinct(subjid, daily_lbdat) %>% 
  nrow()


  

# Create a smaller table whilst working on these graphs
ccp_blood_data <- ccp_data_labs %>%
  select(subjid, cestdat, daily_lbdat, lab_day_since_onset, nl_ratio, 
         daily_wbc_lborres, lymphocyte_count, neutrophil_count, 
         platelet_count, severity2) 

#pivot longer for easier graphing
ccp_blood_data_long <- ccp_blood_data %>%
  pivot_longer(c(nl_ratio, 
                 daily_wbc_lborres, lymphocyte_count, neutrophil_count, 
                 platelet_count), 
               names_to = "assay", 
               values_to = "value")

#add a numeric column with day difference
ccp_blood_data_long <- ccp_blood_data_long %>% 
  mutate(onset_to_lab = as.numeric(lab_day_since_onset))

#check that the above code worked
ccp_blood_data_long %>%
  select(onset_to_lab, lab_day_since_onset)
  head()
```

